### 2019-05-28 models/paper/convergence_paper_models.Rmd
fixing paths

---
title: "Stats for the convergence paper"
author: "Peter Racz"
date: "21/04/2019"
output: html_document
---

```{r echo = F, message = F, warning = F}
knitr::opts_chunk$set(fig.width=10, fig.height=10, fig.path='Figs/', eval=FALSE, echo=TRUE, warning = FALSE, message=FALSE, tidy=TRUE) # eval = FALSE and how

library(brms)
library(lme4)
library(tidyverse)
options(mc.cores=parallel::detectCores())
# options(mc.cores=1)
Sys.setenv(TZ="Europe/Rome") # a szivemben örök tavasz van

try(setwd('../../models/paper'))
try(setwd('~/Github/RaczBecknerHayPierrehumbert2019/models/paper'))
```

These are the models and stats for the convergence paper, going through it by section number.

```{r }
source('../../functions/grabdata.R')

```

## Section 4.2

```{r }
baseline %>% group_by(word) %>% summarise(regular = mean(regular)) %>% filter(word == 'splive') 
```

## Section 4.3

```{r }
espdata %>% select(participant_id,lex_typicality,reg_rate) %>% unique %>% group_by(lex_typicality,reg_rate) %>% summarise(n = n()) 
```

## Section 5

Fits on what people are doing across conditions in test2. No algorithmic learning here.

Our weapon of choice is typially lme4. But it won't converge with full random effects structure. So we turn to brms.

```{r }
test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
test2$reg_rate = factor(test2$reg_rate, levels = c('-40%', 'nc', '+40%'))
test2$baseline_av_z = scales::rescale(test2$baseline_av)
test2$pre_reg_av_z = scales::rescale(test2$pre_reg_av)

fit1 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 | participant_id) + 
    (1 | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit1 = add_criterion(fit1, criterion = 'loo')

save(fit1, file = '../../models/paper/bayes_fit1.rda')

#####################################################

fit2 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit2 = add_criterion(fit2, criterion = 'loo')

save(fit2, file = '../../models/paper/bayes_fit2.rda')

#####################################################

fit3 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit3 = add_criterion(fit3, criterion = 'loo')

# fit3b = add_criterion(fit3b, criterion = 'loo')

save(fit3, file = '../../models/paper/bayes_fit3.rda')

#####################################################

fit4 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z + lex_typicality + reg_rate | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit4 = add_criterion(fit4, criterion = 'loo')

save(fit4, file = '../../models/paper/bayes_fit4.rda')

#####################################################

loo.comparisons = loo_compare(fit1,fit2,fit3,fit4)
save(loo.comparisons, file = '../../models/paper/loo_comparisons.rda')

options(mc.cores=1)

#####################################################
```

```{r }

load('../../models/paper/bayes_fit3.rda')
bayes_factor()

load('../../models/paper/loo_comparisons.rda')
loo.comparisons

# https://arxiv.org/abs/1507.04544 s5.2
```

We definitely want some random slopes (fit1 is worse than the rest). Other than that, the various constellations do not improve the model's expected log pointwise predictive density.

It looks like the model reported in the paper does not ignore too much relevant variation in the random effects.

## Bayes factors!

```{r eval=F}
# random vs atypical

test2b = test2 %>% 
  filter(
    lex_typicality != 'typical'
  )

fit3b = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2b, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15), save_all_pars = T)

fit3c = brm(resp_post_reg ~ 1 + baseline_av_z + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2b, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15), save_all_pars = T)

save(fit3b, file = 'fit3b.rda')
save(fit3c, file = 'fit3c.rda')

```

```{r }
load('fit3b.rda')
load('fit3c.rda')

bf = bayes_factor(fit3c, fit3b)
bf$bf
# time to ask for help.
```

## Section 6

```{r }
baseline$r.baseline_gcm_features = scales::rescale(baseline$baseline_gcm_features, to = c(0,1))
baseline$r.baseline_mgl_features = scales::rescale(baseline$baseline_mgl_features, to = c(0,1))

baseline %>% select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% filter(word == 'splive') %>% unique
baseline %>% select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% filter(word == 'splive') %>% unique

```

## Section 6.3

```{r }
# don't want to load Hmisc, messes with select()
with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .581

with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .586
```

## jumping to Appendix E!

We scale gcm and mgl scores for the regression models. We residualise to see if the baseline gcm and mgl explain variation not explained by the other one.

```{r eval=F}

baseline = baseline %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
  )

baseline_sum = baseline %>% 
  select(word,baseline_gcm_features,c.baseline_gcm_features,baseline_mgl_features,c.baseline_mgl_features) %>% 
  unique

with(baseline_sum, cor(baseline_gcm_features,baseline_mgl_features))
with(baseline_sum, cor(c.baseline_gcm_features,c.baseline_mgl_features))

m1 <- lm(c.baseline_gcm_features ~ c.baseline_mgl_features, data = baseline_sum)
m2 <- lm(c.baseline_mgl_features ~ c.baseline_gcm_features, data = baseline_sum)

summary(m1)
summary(m2)

baseline_sum$gcm_residual <- residuals(m1)
baseline_sum$mgl_residual <- residuals(m2)

baseline_sum = baseline_sum %>% 
  select(gcm_residual,mgl_residual,word)

baseline2 = left_join(baseline,baseline_sum, by = 'word')

m3 = glmer(regular ~ c.baseline_mgl_features + gcm_residual + (1 | subject) + (1|word), data = baseline2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

save(m3, file = '../../models/paper/m3.rda')

m4 = glmer(regular ~ c.baseline_gcm_features + mgl_residual + (1 | subject) + (1|word), data = baseline2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

save(m4, file = '../../models/paper/m4.rda')
```

```{r }

load('../../models/paper/m4.rda')
load('../../models/paper/m3.rda')

summary(m3)
#                         Estimate Std. Error z value Pr(>|z|)    
# (Intercept)             -0.44918    0.11202  -4.010 6.08e-05 ***
# c.baseline_mgl_features  1.27846    0.07525  16.989  < 2e-16 ***
# gcm_residual             1.37767    0.17171   8.023 1.03e-15 ***
summary(m4)
#                         Estimate Std. Error z value Pr(>|z|)    
# (Intercept)             -1.12780    0.14259  -7.909 2.58e-15 ***
# c.baseline_gcm_features  2.29014    0.15126  15.140  < 2e-16 ***
# mgl_residual             0.95005    0.08549  11.113  < 2e-16 ***
```

They do. The abstract generalisations and the word-based comparisons explain non-overlapping domains of variation in participant responses.

## Section 6.4

## jumping to Appendix F; Table 12

```{r }

with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .58
with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .59

with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C']) # .6
with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C']) # .68

with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) # .61
with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C']) # .55
```

## jumping to Appendix F.2

Now we check, for GCM/MGL, if the individual model explains variation beyond the baseline model in the esp test2.

```{r eval=F}
test2 = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features),
    c.individual_gcm_features = scales::rescale(individual_gcm_features),
    c.individual_mgl_features = scales::rescale(individual_mgl_features),
  )

test2$gcm_wf_score_diff <- test2$c.individual_gcm_features - test2$c.baseline_gcm_features
test2$mgl_wf_score_diff <- test2$c.individual_mgl_features - test2$c.baseline_mgl_features

m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m5, file = '../../models/paper/m5.rda')

m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m6, file = '../../models/paper/m6.rda')

# Third

m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m7, file = '../../models/paper/m7.rda')

```

```{r }
load('../../models/paper/m5.rda')
load('../../models/paper/m6.rda')
load('../../models/paper/m7.rda')

summary(m5)
#                  Estimate Std. Error z value Pr(>|z|)    
# (Intercept)              -5.5677     0.3724 -14.949  < 2e-16 ***
# c.baseline_gcm_features   4.3687     0.4249  10.281  < 2e-16 ***
# gcm_wf_score_diff         2.6117     0.4696   5.562 2.67e-08 ***
# pre_reg_av                5.2887     0.6022   8.782  < 2e-16 ***
summary(m6)
#                        Estimate Std. Error z value Pr(>|z|)    
# (Intercept)              -5.0641     0.3374 -15.010   <2e-16 ***
# c.baseline_mgl_features   2.0335     0.2079   9.782   <2e-16 ***
# mgl_wf_score_diff         0.2085     0.1723   1.210    0.226    
# pre_reg_av                7.4492     0.5996  12.425   <2e-16 ***
summary(m7)
#                      Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                -5.6326     0.3364 -16.743  < 2e-16 ***
# c.individual_gcm_features   2.7857     0.3776   7.378 1.61e-13 ***
# c.baseline_mgl_features     1.5998     0.1709   9.361  < 2e-16 ***
# pre_reg_av                  5.4374     0.5842   9.307  < 2e-16 ***
summary(m7)$coef %>% round(3) %>% xtable
# vif.mer(m7)
summary(rePCA(m7))

test2 %>% select(word,moder,verb_baseline_mgl_reg) %>% filter(moder == 'KEPT' & verb_baseline_mgl_reg == 1) %>% unique %>% arrange(word) %>% select(word) %>% pull %>% length
# 28, not 57
```

It looks like the best model has the predictions of the individual gcm and the baseline mgl.

## Marginal and conditional r squared for the two models

```{r eval=F}
load('esp_fit.rda')
formula(fit)
MuMIn::r.squaredGLMM(fit) # 0.27 0.45
MuMIn::r.squaredGLMM(m7) # 0.22 0.47 
test2$m7.pred = predict(m7, type = 'response')
with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) # .84
```

## Figure 4

```{r }
# baseline, post-test, type
# type: means, gcm, mgl
b1 = baseline %>% 
  group_by(
    word,
    baseline_gcm_features,
    baseline_mgl_features
    ) %>% 
  summarise(reg.av = mean(regular)) %>% 
  ungroup()

b1 = b1 %>% 
  mutate(
    baseline_gcm_features = scales::rescale(baseline_gcm_features, to = c(0,1)),
    baseline_mgl_features = scales::rescale(baseline_mgl_features, to = c(0,1)),
    reg.av = scales::rescale(reg.av, to = c(0,1))
  )

t2 = test2 %>%
  mutate(lex_typicality2 = 
           case_when(
             lex_typicality == 'atypical' ~ 'reversed',
             lex_typicality == 'typical' ~ 'typical',
             lex_typicality == 'random' ~ 'random',
           ),
         lex_typicality2 = factor(lex_typicality2, levels = c('typical', 'reversed', 'random'))
         ) %>% 
  group_by(
    word, 
    lex_typicality2
  ) %>% 
  summarise(
    post.reg.av = mean(resp_post_reg),
    individual_gcm_features = mean(individual_gcm_features),
    individual_mgl_features = mean(individual_mgl_features),
  ) %>% 
  ungroup() %>% 
  mutate(
    individual_gcm_features = scales::rescale(individual_gcm_features, to = c(0,1)),
    individual_mgl_features = scales::rescale(individual_mgl_features, to = c(0,1)),
    post.reg.av = scales::rescale(post.reg.av, to = c(0,1))
  )

t3 = t2 %>% left_join(b1, by = 'word') %>% 
  rename(
    individual.mgl = individual_mgl_features,
    individual.gcm = individual_gcm_features,
    baseline.mgl = baseline_mgl_features,
    baseline.gcm = baseline_gcm_features
  )

t3la = t3 %>% 
  select(
    word,
    lex_typicality2,
    baseline.gcm,
    baseline.mgl,
    reg.av
  ) %>% 
  rename(
    GCM = baseline.gcm,
    MGL = baseline.mgl,
    participants = reg.av
  ) %>% 
  gather(type, baseline.averages, -word, -lex_typicality2)

t3lb = t3 %>% 
  select(
    word,
    lex_typicality2,
    individual.gcm,
    individual.mgl,
    post.reg.av
  ) %>% 
  rename(
    GCM = individual.gcm,
    MGL = individual.mgl,
    participants = post.reg.av
  ) %>% 
  gather(type, individual.averages, -word, -lex_typicality2)

t4 = full_join(t3la,t3lb, by = c("word", "lex_typicality2", "type"))

t4$type = factor(t4$type, levels = c('participants', 'GCM', 'MGL'))

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline average') +
  ylab('verb ESP average')
# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
ggsave('fig4_2019.pdf', width = 12.5, height = 4)

```