### 2020-01-28 models/paper/convergence_paper_models.Rmd
in progress

---
title: "Stats for the convergence paper"
author: "Peter Racz"
date: "21/04/2019"
output: html_document
---

```{r echo = F, message = F, warning = F}
knitr::opts_chunk$set(fig.width=10, fig.height=10, fig.path='Figs/', eval=FALSE, echo=TRUE, warning = FALSE, message=FALSE, tidy=TRUE) # eval = FALSE and how

library(brms)
library(lme4)
library(merTools)
library(tidyverse)
options(mc.cores=parallel::detectCores())
# options(mc.cores=1)
Sys.setenv(TZ="Europe/Rome") # a szivemben örök tavasz van

try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
```

These are the models and stats reported in the main text of the convergence paper. We go through the main text by section number and replicate the relevant statistics / figures / models here.

This file references both the main text and the supplementary information. To make things slightly more complicated, the appendix of the main text and the SI overlap. And section lettering is different in the appendix of the main text and the si. Here, we use lettering that matches the si.

```{r }
source('functions/grabdata.R')

espdata[espdata$lex_typicality == 'atypical',]$lex_typicality = 'reversed'

test2[test2$lex_typicality == 'atypical',]$lex_typicality = 'reversed'

```

## Section 4.2

```{r }
baseline %>% 
  group_by(word) %>% 
  summarise(regular = mean(regular)) %>% 
  filter(word == 'fim') 
```

## Section 4.3

```{r }
espdata %>% 
  select(participant_id,lex_typicality,reg_rate) %>% 
  unique %>% 
  group_by(lex_typicality,reg_rate) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = reg_rate, values_from = n)
```

## Section 5

Fits on what people are doing across conditions in test2. No algorithmic learning here.

Our weapon of choice is typially lme4. But it won't converge with full random effects structure. So we turn to brms. We assume flat priors and a bernoulli error distribution, which is preferred to the binomial distribution in brms for binary outcomes. We compare models using leave-one-out cross-validation.

The Bayesian models are equivalent to the frequentist models reported in the paper, at least conceptually, if not in many other ways.

```{r }
fit1 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 | participant_id) + 
    (1 | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit1 = add_criterion(fit1, criterion = 'loo')

save(fit1, file = '../../models/paper/bayes_fit1.rda')

#####################################################

fit2 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit2 = add_criterion(fit2, criterion = 'loo')

save(fit2, file = '../../models/paper/bayes_fit2.rda')

#####################################################

fit3 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit3 = add_criterion(fit3, criterion = 'loo')

# fit3b = add_criterion(fit3b, criterion = 'loo')

save(fit3, file = '../../models/paper/bayes_fit3.rda')

#####################################################

fit4 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z + lex_typicality + reg_rate | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit4 = add_criterion(fit4, criterion = 'loo')

save(fit4, file = '../../models/paper/bayes_fit4.rda')

#####################################################

loo.comparisons = loo_compare(fit1,fit2,fit3,fit4)
save(loo.comparisons, file = '../../models/paper/loo_comparisons.rda')

options(mc.cores=1)

#####################################################
```

```{r }

load('models/paper/bayes_fit3.rda')
bayes_factor()

load('../../models/paper/loo_comparisons.rda')
loo.comparisons
```

What do we learn from LOO comparisons?

Fit1 is worse than the rest, so random slopes do improve model fit. Building up the random effect structure for the word random intercept does not improve the model's expected log pointwise predictive density.

It looks like the model reported in the paper does not ignore too much relevant variation in the random effects.

## Bayes factors

We could not reject the null on the absence of difference between our typical and atypical (reversed) conditions. Upon reviewer suggestion, we used Bayes factors to test whether this is due to the lack of evidence.

```{r eval=F}
# random vs atypical

test2b = test2 %>% 
  filter(
    lex_typicality != 'typical'
  )

fit_power1 = glmer(resp_post_reg ~ 1 + baseline_av_c * lex_typicality + pre_reg_av_c + reg_rate +
    (1 + baseline_av_c | participant_id) + 
    (1 | word), 
  data = test2b, 
  family = binomial ( link = logit ),
  control=glmerControl(optimizer="bobyqa")
  )

fit_power2 = glmer(resp_post_reg ~ 1 + baseline_av_c + lex_typicality + pre_reg_av_c + reg_rate +
    (1 + baseline_av_c | participant_id) + 
    (1 | word), 
  data = test2b, 
  family = binomial ( link = logit ),
  control=glmerControl(optimizer="bobyqa")
  )

# summary(fit_power1)
# summary(fit_power2)

exp((BIC(fit_power2) - BIC(fit_power1))/2)
# 0.01179529
# it is unlikely that there is a difference here

```

The lme4 fit reported in the paper is here:

```{r }
# the frequentist fit reported in the paper
# watch out for factor levels
test2$verb_baseline_average = test2$baseline_av_c
test2$bot_lexical_typicality = test2$lexical_typicality
test2$participant_prereg_average = test2$pre_reg_av_c
test2$bot_regularisation_rate = test2$regularisation_rate

lme.fit2 = glmer(resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
    (1 + baseline_av_c | participant_id) + 
    (1  | word), 
  data = test2, 
  family = binomial (link = logit),
  control=glmerControl(optimizer="bobyqa")
)
summary(lme.fit2)$coef

plot(effects::effect('bot_regularisation_rate', lme.fit2), main = '', xlab = 'regularisation shift', ylab = 'regular response in post-test')
plot(effects::effect('verb_baseline_average:bot_lexical_typicality', lme.fit2), multiline = T, lines=list(lty=c(1,2,3)), main = '', xlab = 'verb baseline mean', ylab = 'regular response in post-test')

# https://arxiv.org/abs/1507.04544 s5.2
```

Here is a simple confint plot using Wald estimates.

```{r }
t1 = broom::tidy(lme.fit2)

# we get some wald confidence intervals

confints = confint(lme.fit2, method = 'Wald') %>% broom::tidy()
confints = confints[-1,]
names(confints) = c('term', '2.5%', '97.5%')
t1 = inner_join(t1, confints) %>% 
  dplyr::select(-p.value,-group)

t1 %>% 
  mutate(
    row_id = 9:1, # watch this
  ) %>% 
  mutate(
    term2 = 
    case_when(
      term == 'verb_baseline_average' ~ 'verb baseline average',
      term == 'bot_lexical_typicalityrandom' ~ 'bot lexical typicality = random',
      term == 'bot_lexical_typicalityatypical' ~ 'bot lexical typicality = atypical',
      term == 'participant_prereg_average' ~ 'participant prereg average',
      term == 'bot_regularisation_rate-40%' ~ 'bot regularisation rate = -40%',
      term == 'bot_regularisation_rate+40%' ~ 'bot regularisation rate = +40%',
      term == 'verb_baseline_average:bot_lexical_typicalityrandom' ~ 'verb baseline average : bot lexical typicality = random',
      term == 'verb_baseline_average:bot_lexical_typicalityatypical' ~ 'verb baseline average : bot lexical typicality = reversed',
      term == '(Intercept)' ~ 'Intercept'
      )
    ) %>% 
  ggplot(aes(x = term2 %>% reorder(row_id), y = estimate)) +
  geom_point() +
  geom_linerange(aes(x = term2, ymin = `2.5%`, ymax = `97.5%`)) +
  geom_hline(aes(yintercept = 0), linetype = 'dashed') +
  xlab('') +
  coord_flip() +
  ggtitle('Estimates and 95% confidence intervals in model for test results')
ggsave('terms_figure.pdf', width = 9, height = 5)

```

Let's have a few predicted intervals.

```{r }
lme.fit2.preds = predictInterval(merMod = lme.fit2, level = 0.95, n.sims = 1000, stat = 'median', type = 'probability')

test2c = cbind(test2,lme.fit2.preds) 
test2c %>%   
  group_by(
    reg_rate
  ) %>% 
  summarise(
    fit = mean(fit),
    upr = mean(upr),
    lwr = mean(lwr)
  ) %>% 
  ggplot(aes(x = reg_rate, y = fit)) +
  geom_point() +
  geom_linerange(aes(ymin = lwr, ymax = upr))

test2 %>% 
  mutate(
    reg_rate2 = factor(reg_rate, levels = c('-40%', 'nc', '+40%'))
  ) %>% 
  ggplot(aes(x = reg_rate2, y = post_reg_av, fill = reg_rate2)) +
  geom_jitter(aes(colour = reg_rate2), width = 0.3) +
    geom_violin(aes(alpha = 0.5)) +
    geom_boxplot(width = 0.2) +
    # stat_summary(fun.y=mean, geom="point", shape=16, size=4) +
    scale_fill_brewer(palette="Set2") +
    scale_colour_brewer(palette="Set2") +
    theme(
      text = element_text(size=30), 
      # axis.text.x = element_text(angle=90, vjust=0.5), 
      # axis.title.x=element_blank(), 
      legend.position = 'none'
      ) +
    xlab('bot regularisation shift') +
    ylab('participant post-test mean regularisation') +
    scale_x_discrete(labels=c('-40%', 'no change', '+40%'))
ggsave('reg_shift_av.pdf', width = 9, height = 6)

test2 %>% 
  group_by(
    word,
    baseline_av,
    lex_typicality
  ) %>% 
  summarise(
    mean = mean(resp_post_reg)
  ) %>% 
  ggplot(aes(x = baseline_av, y = mean, colour = lex_typicality, label = word)) +
  geom_label() +
  geom_smooth(method = 'lm', se = F) +
  scale_colour_brewer(palette="Dark2") +
    theme(
      text = element_text(size=30), 
      # axis.text.x = element_text(angle=90, vjust=0.5), 
      # axis.title.x=element_blank(), 
      legend.position = 'none'
      ) +
    xlab('verb baseline mean regularisation') +
    ylab('verb post-test mean regularisation')
ggsave('lex_typ_av.pdf', width = 9, height = 9)

sum1 = confint(lme.fit2, method = 'Wald') %>% data.frame
sum2 = summary(lme.fit2)$coef %>% data.frame
sum1$Term = rownames(sum1)
sum2$Term = rownames(sum2)
sum3 = left_join(sum2, sum1) %>% 
  rename(`Std error` = Std..Error,
    `z value` = z.value,
    `2.5%` = X2.5..,
    `97.5%` = X97.5..
    )
sum3 %>% dplyr::select(Term, Estimate, `Std error`, `z value`, `2.5%`, `97.5%`) %>% knitr::kable(digits = 2)

```

## Section 6

```{r }
baseline$r.baseline_gcm_features = scales::rescale(baseline$baseline_gcm_features, to = c(0,1))
baseline$r.baseline_mgl_features = scales::rescale(baseline$baseline_mgl_features, to = c(0,1))

baseline %>% 
  select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% 
  filter(word == 'splive') %>% 
  unique

baseline %>% 
  select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% 
  filter(word == 'splive') %>% 
  unique

```

## Section 6.3

```{r }
# don't want to load Hmisc, messes with select()
with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C'])

with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) 
```

## jumping to SI section E!

We scale gcm and mgl scores for the regression models. We residualise to see if the baseline gcm and mgl explain variation not explained by the other one.

```{r eval=F}

baseline = baseline %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
  )

baseline_sum = baseline %>% 
  select(word,baseline_gcm_features,c.baseline_gcm_features,baseline_mgl_features,c.baseline_mgl_features) %>% 
  unique

with(baseline_sum, cor(baseline_gcm_features,baseline_mgl_features))
with(baseline_sum, cor(c.baseline_gcm_features,c.baseline_mgl_features))

m1 <- lm(c.baseline_gcm_features ~ c.baseline_mgl_features, data = baseline_sum)
m2 <- lm(c.baseline_mgl_features ~ c.baseline_gcm_features, data = baseline_sum)

summary(m1)
summary(m2)

baseline_sum$gcm_residual <- residuals(m1)
baseline_sum$mgl_residual <- residuals(m2)

baseline_sum = baseline_sum %>% 
  select(gcm_residual,mgl_residual,word)

baseline2 = left_join(baseline,baseline_sum, by = 'word')

m3 = glmer(regular ~ c.baseline_mgl_features + gcm_residual + (1 | subject) + (1|word), data = baseline2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

save(m3, file = 'models/paper/m3.rda')

m4 = glmer(regular ~ c.baseline_gcm_features + mgl_residual + (1 | subject) + (1|word), data = baseline2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

save(m4, file = 'models/paper/m4.rda')
```

```{r }

load('models/paper/m4.rda')
load('models/paper/m3.rda')

summary(m3)

summary(m4)

```

They do. The abstract generalisations and the word-based comparisons explain non-overlapping domains of variation in participant responses.

## Section 6.4

## jumping to SI section F; Table 12

```{r }

# first method

with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) 
with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C'])

with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C'])
with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C'])

with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) 
with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C'])
```

## jumping to SI section F.2

Now we check, for GCM/MGL, if the individual model explains variation beyond the baseline model in the esp test2.

```{r eval=F}

# second method

test2 = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features),
    c.individual_gcm_features = scales::rescale(individual_gcm_features),
    c.individual_mgl_features = scales::rescale(individual_mgl_features),
  )

test2$gcm_wf_score_diff <- test2$c.individual_gcm_features - test2$c.baseline_gcm_features
test2$mgl_wf_score_diff <- test2$c.individual_mgl_features - test2$c.baseline_mgl_features

m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m5, file = 'models/paper/m5.rda')

m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m6, file = 'models/paper/m6.rda')

# third method

m7a <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + c.individual_mgl_features + c.baseline_gcm_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

# summary(m7a)

m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m7, file = 'models/paper/m7.rda')

```

```{r }
load('models/paper/m5.rda')
load('models/paper/m6.rda')
load('models/paper/m7.rda')

summary(m5)
summary(m6)
summary(m7)
summary(m7)$coef %>% round(3) %>% xtable
# vif.mer(m7)
summary(rePCA(m7))

test2 %>% 
  select(word,moder,verb_baseline_mgl_reg) %>% 
  filter(moder == 'KEPT' & verb_baseline_mgl_reg == 1) %>% 
  unique %>% 
  arrange(word) %>% 
  select(word) %>% 
  pull %>% 
  length
# 28, not 57
```

It looks like the best model has the predictions of the individual gcm and the baseline mgl.

## Marginal and conditional r squared for the two models

```{r eval=F}
load('models/paper/esp_fit.rda')
formula(fit)
MuMIn::r.squaredGLMM(fit) # 0.27 0.45
MuMIn::r.squaredGLMM(m7) # 0.22 0.47 
test2$m7.pred = predict(m7, type = 'response')
with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) 
```

## Figure 4

```{r }
# baseline, post-test, type
# type: means, gcm, mgl
b1 = baseline %>% 
  group_by(
    word,
    baseline_gcm_features,
    baseline_mgl_features
    ) %>% 
  summarise(reg.av = mean(regular)) %>% 
  ungroup()

b1 = b1 %>% 
  mutate(
    baseline_gcm_features = scales::rescale(baseline_gcm_features, to = c(0,1)),
    baseline_mgl_features = scales::rescale(baseline_mgl_features, to = c(0,1)),
    reg.av = scales::rescale(reg.av, to = c(0,1))
  )

t2 = test2 %>%
  mutate(lex_typicality2 = 
           case_when(
             lex_typicality == 'atypical' ~ 'reversed',
             lex_typicality == 'reversed' ~ 'reversed',
             lex_typicality == 'typical' ~ 'typical',
             lex_typicality == 'random' ~ 'random',
           ),
         lex_typicality2 = factor(lex_typicality2, levels = c('typical', 'reversed', 'random'))
         ) %>% 
  group_by(
    word, 
    lex_typicality2
  ) %>% 
  summarise(
    post.reg.av = mean(resp_post_reg),
    individual_gcm_features = mean(individual_gcm_features),
    individual_mgl_features = mean(individual_mgl_features),
  ) %>% 
  ungroup() %>% 
  mutate(
    individual_gcm_features = scales::rescale(individual_gcm_features, to = c(0,1)),
    individual_mgl_features = scales::rescale(individual_mgl_features, to = c(0,1)),
    post.reg.av = scales::rescale(post.reg.av, to = c(0,1))
  )

t3 = t2 %>% left_join(b1, by = 'word') %>% 
  rename(
    individual.mgl = individual_mgl_features,
    individual.gcm = individual_gcm_features,
    baseline.mgl = baseline_mgl_features,
    baseline.gcm = baseline_gcm_features
  )

t3la = t3 %>% 
  dplyr::select(
    word,
    lex_typicality2,
    baseline.gcm,
    baseline.mgl,
    reg.av
  ) %>% 
  rename(
    GCM = baseline.gcm,
    MGL = baseline.mgl,
    participants = reg.av
  ) %>% 
  gather(type, baseline.averages, -word, -lex_typicality2)

t3lb = t3 %>% 
  dplyr::select(
    word,
    lex_typicality2,
    individual.gcm,
    individual.mgl,
    post.reg.av
  ) %>% 
  rename(
    GCM = individual.gcm,
    MGL = individual.mgl,
    participants = post.reg.av
  ) %>% 
  gather(type, individual.averages, -word, -lex_typicality2)

t4 = full_join(t3la,t3lb, by = c("word", "lex_typicality2", "type"))

t4$type = factor(t4$type, levels = c('participants', 'GCM', 'MGL'))

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2, linetype = lex_typicality2)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type) +
  scale_shape_discrete(name="Lexical\ntypicality") +
  # scale_colour_discrete(name="Lexical\ntypicality") +
  xlab('verb baseline regularization') +
  ylab('verb ESP regularization') +
  guides(linetype = F) +
  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
ggsave('fig4_2020.pdf', width = 12.5, height = 4)

```